---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(PrepR)
```

## PrepR 

<!-- badges: start -->
[![R build status](https://github.com/UBC-MDS/PrepR/workflows/R-CMD-check/badge.svg)](https://github.com/UBC-MDS/PrepR/actions) [![codecov](https://codecov.io/gh/UBC-MDS/PrepR/branch/master/graph/badge.svg)](https://codecov.io/gh/UBC-MDS/PrepR)
<!-- badges: end -->

### Package Summary

`PrepR` is a package for R to help preprocessing in machine learning tasks. 
There are certain repetitive tasks  that come up often when doing a machine learning project and this package aims to alleviate those chores.
Some of the issues that come up regularly are: finding the types of each column in a dataframe, splitting the data (whether into train/test sets or train/test/validation sets, one-hot encoding,  and scaling features.
This package will help with all of those tasks.

Please find detailed descriptions of the functions in this [vignette](https://ubc-mds.github.io/PrepR/articles/my-vignette.html).

### Installation:

You can install the released version of PrepR from [CRAN](https://CRAN.R-project.org) with:

``` r
# install.packages("PrepR")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("UBC-MDS/PrepR")
```

### Features
This package has the following features:

- `train_valid_test_split`: This function splits the data set into train, validation, and test sets.

- `data_type`: This function identifies data types for each column/feature. It returns one dataframe for each type of data.

- `one-hot`: This function performs one-hot encoding on the categorical features and returns a dataframe with sensible column names.

- `scaler`: This function performs standard scaling on the numerical features. 

### Dependencies

R version 3.6.1 and R packages:

- dplyr
- gensvm
- tibble
- nlme
- magrittr

### Usage

`library(PrepR)`

**Identify features of different data types**  

`data_type(my_data)$num`

`data_type(my_data)$cat`  

Examples: 
```{r example}
fruits_df <- data.frame(fruits = c("apple", "pear", "apple", "banana", "orange"), count = c(1L, 2L, 3L, 4L, 5L), price = c(2, 3.4, 0.5, 7, 8))
fruits_y <- data.frame(target = c(1, 2, 3, 4, 5))

fruits_df
```


```{r example_datatype}
data_type(fruits_df)$num
```


**Train, validation, and test split**  

`train_valid_test_split(X, y, test_size, valid_size, train_size, stratify, random_state, shuffle)`

Examples: 
```{r example_split}
train_valid_test_split(fruits_df, fruits_y)$x_train
```
  
  
**One-hot encode features of categorical type**  

`onehot(my_data)`  

Examples: 
```{r example_onehot}
onehot(data_type(fruits_df)$cat)
```


**Standard Scaling of categorical features**  

`scaler(x_train, x_test, colnames)$x_train`

`scaler(x_train, x_test, colnames)$x_test`  

Examples: 
```{r example_scaling, warning = FALSE}
num_var <- data_type(fruits_df)$num

scaler(num_var, num_var, num_var, c("count", "price"))$x_train
```

### Our package in the R ecosystem

There are no packages that do everything that this one does, but there are packages for machine learning in R that this package will make use of. 
The `caret` package in R does some of these preprocessing steps, such as train/test split. 
It does not, however, have a function that takes a dataframe and returns multiple dataframes based on their column type; 
this is the issue that this package's `data_type` function seeks to solve.
However, it does not seem that there is an option for having a validation set in `caret`'s version of train/test split.
This is an option that would be useful to anyone doing machine learning.
The `caret` package also does one-hot encoding with the function `dummyVars`, though `onehot()` in the PrepR package is more intuitive:
its function does not return meaningful, human-readable column names.
It also removes one column by default, which is better for fast computation, but worse for human-readability.

Overall, this package fits in well with the R ecosystem and helps make machine learning a little easier. 


### Documentation
The official documentation is hosted on Read the Docs: <https://PrepR.readthedocs.io/en/latest/>

### Credits
This package was created using the [Whole Game Chapter](https://r-pkgs.org/whole-game.html) from the [R Packages](https://r-pkgs.org/) handbook by Hadley Wickham
