---
title: "PrepR-vignette"
output: rmarkdown::html_vignette
author: Chimaobi Amadi, Matthew Connell, Jasmine Qin, George Thio
date: March 13, 2020
vignette: >
  %\VignetteIndexEntry{PrepR-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`PrepR` is a package for R to help preprocessing in machine learning tasks. There are four main functions included in this package that fit into four important steps in general data preprocessing pipeline.

```{r setup}
library(PrepR)
```

## Load Data
To explore functions in `PrepR`, let's first load some toy data. Here we have one categorical data and two numerical data

```{r load_data}
fruits_df <- data.frame(fruits = c("apple", "pear", "apple", "banana", "orange"), count = c(1L, 2L, 3L, 4L, 5L), price = c(2, 3.4, 0.5, 7, 8))
fruits_y <- data.frame(target = c(1, 2, 3, 4, 5))
```

## Train Test Split
The first thing to do in a machine learning pipeline is to split the data in to train, validation, and test datasets. We will demonstrating this using the function `train_valid_test_split`.

```{r train_valid_test_split}
x_train <- train_valid_test_split(fruits_df, fruits_y)$x_train
x_valid <- train_valid_test_split(fruits_df, fruits_y)$x_valid
x_test <- train_valid_test_split(fruits_df, fruits_y)$x_test
y_train <- train_valid_test_split(fruits_df, fruits_y)$y_train
y_valid <- train_valid_test_split(fruits_df, fruits_y)$y_valid
y_test <- train_valid_test_split(fruits_df, fruits_y)$y_test
x_train
```

## Distinguish Data Types
Next we will need to know variable types of the columns in order to do further transformation or analysis on them. Let's try using our `data_type` function to split categorical and numeric data.

```{r data_type}
numeric_variables_train <- PrepR::data_type(x_train)$num
numeric_variables_valid <- PrepR::data_type(x_valid)$num
numeric_variables_test <- PrepR::data_type(x_test)$num
categorical_variables <- PrepR::data_type(x_train)$cat
categorical_variables
```

## One Hot Encode
One hot encoding is a powerful transformation that allows us to work with categorical data. Let's take a look at what our `onehot` does.

```{r one-hot}
onehot(categorical_variables)
```

## Numeric Scaling
Now we're left with our numeric variables only. We usually want to normalize our numeric variables for better modelling performance, for example for k-Nearest Neighbours algorithm. The `scaler` function in our package will do this for you.

```{r scaling}
scaler(numeric_variables_train, numeric_variables_valid, numeric_variables_test, c("count", "price"))$X_train
```


